---
title: "Plot_Stats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot_Stats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(G6)
sustainability
```

## Plot and stats

The purpose of the function "Plot and Stats" is to compute the "sustainability score" that we defined as the difference between Biocapacity per Capita of a country and the Ecological Footrpint of Consumption per Capita.

If the score is negative, it means that the country consume more resources than the land is able to provide to the country. 

The data are retrieved from the NFA 2019 public_data file that we can find in the National Futures Association website. 

### Functioning 

From the dataset, we 

1. Filter the dataset to get only the BiocapPerCap & EFConsPerCap data, 
1. Group these data by country and by year, 
1. Create the sustainability column

After these steps, we are able to plot the Sustainability by country and by year, to have a visual of the sustainability value of each country. 

### Shiny App
